Modeling Human Interactions in Collaborative Interactive Evolutionary Computation

### 2017-02-02 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
EvoSpace-I

\section{HC  IEC Framework}
% Comment: 
% Maybe in this case we should use the term software framework or architecture. 
The general goal of this research is to develop a human-centered \cite{gasson2003human} 
software framework for interactive evolutionary computation (IEC). 
The framework is a reusable architectural design together with an implementation \cite{campbell1991choices}, 
providing generalized components to developers of IEC systems. In the description we deliberately use the more general 
term Human instead of User, because humans are not always playing the role of users in software engineering sense, they
could just be interacting naturally with their environment and with out knowing they are implicitly
evaluating a possible solution. The framework includes components that can be refined to increase
participation and also to minimize the amount of evaluations needed for the evolutionary process in a given IEC application. Software frameworks often have a vison or a way of thinking \cite{carneiro2010introducing} guiding their
design. Before entering to details the main design decisions of the framework are explained next:

@@

\item {\bf Interaction is a stream of actions}
  Real time processing of users' actions could be needed for certain applications when data is 
  captured by sensors, or directly captured as user input.
  Graph based models are currently used in social networks to keep track  
  of user interactions with media objects, places, and other users. Users of 
  social networks (for instance the Facebook Graph) are accustomed to express these 
  complex relationships in sentences such as: ``John and Ann eating breakfast at Tony's''. 
  Other example is the W3C Activity Streams 2.0 specification used for representing activities 
  common in social web applications \cite{json:streams}. An important component of the 
  framework is the stream of user-activities.
\end{itemize}

The human centered framework consists of three high level components depicted
in figure \ref{fig:uc_framework}:

\begin{itemize}
  \item {\bf Interactive Evolutionary Computational system} 
  This is the real world system that we are going to represent in the model, 
  it consist of human users and their interactions with one or more phenotypes
  from
  the population. There are many ways in which humans could interact 
  with phenotypes, this could be through a mobile device, a user interface 
  or by interacting with real world objects generated as part of the algorithm 
  \cite{de2014artists,de2013unplugging}. 
  There is also the possibility that fitness or even part of the search 
  to be done by devices lent by the human \cite{DBLP:conf/gecco/MereloCGCRV16}.
  The information gathered through the interaction is the primary concern
  as it will guide the search.

  \item {\bf IEC Data Model}
  A data model is used to describe the IEC system prior to a physical 
  implementation.  Depending on the domain several techniques can be used.
  Parts of the system could be better described by an Entity relationship 
  modeling to be used in a relational database or a class digram for a 
  key-value store.

  A graph is proposed for modeling the social network of users 
  and their interaction with the population. When implemented a graph database 
  will be the back-end of the system. This  backend will then be
  used by the EA and the user interface, in a Model-View-Controller architecture.

  \item {\bf Evolutionary Algorithm} 
  The EA algorithm queries the graph model, and then updates those nodes that 
  are part of the population. The EA could also be done with the help of
  humans, for instance in the XY project \cite{de2013unplugging} artists physically created
  the new population.  
\end{itemize}

\section{EvoSpace-Interactive}
\label{sec:evospace-i}
\begin{figure}[!t]
    \centering
        \includegraphics[width=3.5in]{img/UI_ed01.png}
    \caption{User interface of an EvoSpace-Interactive application.}
    \label{fig:web}
\end{figure}

As a case study, a IEA was developed using the 
EvoSpace-Interactive (ES-I) platform \cite{garcia2013evospace}
A brief description of the application is presented next, focusing
on the data elements that where ported to the graph model.

@@

\subsection{Collaboration}
\label{sec:col}
After entering the web application by using their Facebook account,
users can collaborate with their Facebook friends, 
sharing those phenotypes they like, or by taking phenotypes
from their friend's collections by using the web interface depicted 
in Figure \ref{fig:web}.
At the top left corner a list of Facebook friends is presented
to encourage users to interact with the system. In the central 
\emph{ Wall } area, a phenotype sampled from the population that is
being evolved via the evolutionary algorithm 
is shown to the user.
Here, the user can interact with the system in two ways.
First, he can assign a quality assessment to the phenotype using
a five star rating system or,
additionally, a user can choose to add an image to one of their \emph{Collections}.
A collection is a special folder that stores those phenotypes a user likes and wishes
to save. After the user finishes interacting with the phenotype
on the Wall, he can choose to retrieve a new one from the population.
At the left hand side, the web page shows the \emph{Collections} section.
The user can create several collections, to group and organize his favorite 
artifacts. Moreover, users can browse the content of each collection and from
there share images through their social network.
When a user hovers the cursor over a phenotypes a pop-up pane shows how many users have
liked the phenotype. The pane also includes a link to the phenotype's 
details, the parents, genetic operators that created it, and genealogy
information. This makes the assignment of fitness through the rating
system a {\em social} activity, pursuing the objective of this work,
which is to increase engagement of the users.

\begin{figure*}[!t]
    \centering
        \includegraphics[width=3.5in]{img/framework.png}
    \caption{IEC Human-centered framework.}
    \label{fig:uc_framework}
\end{figure*}

  \subsection{Interfaces}
  \subsection{Data Models}
    \subsubsection{EvoSpace}
    \subsubsection{Graph}
    \subsubsection{PostgreSQL}
  \subsection{Evolution}

\section{Case Studies}
\subsection{EvoDraw}
\subsection{EvoDraw-Kinect}
\subsection{XYZ-Unplugged}

>>>>

\begin{figure*}[!t]
    \centering
        \includegraphics[width=4.5in]{img/framework.eps}
    \caption{IEC Human-centered framework.}
    \label{fig:uc_framework}
\end{figure*}

\section{Human-Centered IEC Framework}
% Comment: 
% Maybe in this case we should use the term software framework or architecture. 
The general goal of this research is to develop a human-centered \cite{gasson2003human} 
software framework for interactive evolutionary computation (IEC). 
The framework is a reusable architectural design together with an implementation \cite{campbell1991choices}, 
providing generalized components to developers of IEC systems. In the description we deliberately use the more general 
term Human instead of User, because humans are not always playing the role of users in software engineering sense, they
could just be interacting naturally with their environment and with out knowing they are implicitly
evaluating a possible solution. The framework includes components that can be refined to increase
participation and also to minimize the amount of evaluations needed for the evolutionary process in a given IEC application. Software frameworks often have a vison or a way of thinking \cite{carneiro2010introducing} guiding their
design. Before entering to details the main design decisions of the framework are explained next:

@@

\item {\bf Interaction is a stream of actions}
  Real time processing of users' actions could be needed for certain applications when data is 
  captured by sensors, or directly captured as user input. For example, social networks keep track of 
  users' interactions with other users, media objects and places. Users of 
  social networks (for instance the Facebook Graph) are accustomed to express these 
  complex relationships in sentences such as: ``John and Ann eating breakfast at Tony's''. 
  Other example is the W3C Activity Streams 2.0 specification used for representing activities 
  common in social web applications \cite{json:streams}. 
\end{itemize}

The human centered IEC architecture consists of three high level components depicted
in figure \ref{fig:uc_framework}:

\begin{itemize}
  \item {\bf Interactive Evolutionary Computational system} 
  This is the real world system that we are going to represent in the data model, 
  it consist of human users and their interactions with one or more phenotypes
  from the population. There are many ways in which humans could interact 
  with phenotypes. The interaction consists of a set of actions and 
  takes place in a certain context. For example, through a mobile device  
  or by interacting with real world objects 
  \cite{de2014artists,de2013unplugging}. 
  There is also the possibility that fitness or even part of the search 
  is done by devices lent by humans \cite{DBLP:conf/gecco/MereloCGCRV16}.
  The information gathered through the interaction is the primary concern
  as it will guide the search.

  \item {\bf IEC Data Model}
  A data model is used to describe the IEC system prior to a physical 
  implementation.  Depending on the domain several techniques can be used.
  Parts of the system could be better described by an Entity relationship 
  modeling to be used in a relational database or a class digram for a 
  key-value store. A graph is proposed for modeling the social network of users 
  and their interaction with the population. When implemented a graph database 
  will be the back-end of the system.

  \item {\bf Evolutionary Algorithm} 
  The EA algorithm interacts with the data model. The EA could also be done with the help of
  humans, for instance in the XY project \cite{de2013unplugging} artists actually painted
  the solutions of the new population.  
\end{itemize}

\section{Data Model Implementation}
The core of the framework is the data model because both the IEC system and the EA will 
depend on the application. In this section implementation details of the Data Model are presented. 
  \subsection{EvoSpace}
  \subsection{Graph}
  \subsection{PostgreSQL}

  \subsection{Interfaces}

  \subsection{Evolution}

\section{Data Model Implementation}

\section{Case Studies}
\label{sec:evospace-i}
\begin{figure}[!t]
    \centering
        \includegraphics[width=3.5in]{img/interface.png}
    \caption{User interface of the EvoDrawings application.}
    \label{fig:web}
\end{figure}

\subsection{EvoDraw}
As a case study, a IEA was developed using the 
EvoSpace-Interactive (ES-I) platform \cite{garcia2013evospace}
A brief description of the application is presented next, focusing
on the data elements that where ported to the graph model.

@@

\subsection{Collaboration}
\label{sec:col}
After entering the web application by using their Facebook account,
users can collaborate with their Facebook friends, 
sharing those phenotypes they like, or by taking phenotypes
from their friend's collections by using the web interface depicted 
in Figure \ref{fig:web}.
At the top left corner a list of Facebook friends is presented
to encourage users to interact with the system. In the central 
\emph{ Wall } area, a phenotype sampled from the population that is
being evolved via the evolutionary algorithm 
is shown to the user.
Here, the user can interact with the system in two ways.
First, he can assign a quality assessment to the phenotype using
a five star rating system or,
additionally, a user can choose to add an image to one of their \emph{Collections}.
A collection is a special folder that stores those phenotypes a user likes and wishes
to save. After the user finishes interacting with the phenotype
on the Wall, he can choose to retrieve a new one from the population.
At the left hand side, the web page shows the \emph{Collections} section.
The user can create several collections, to group and organize his favorite 
artifacts. Moreover, users can browse the content of each collection and from
there share images through their social network.
When a user hovers the cursor over a phenotypes a pop-up pane shows how many users have
liked the phenotype. The pane also includes a link to the phenotype's 
details, the parents, genetic operators that created it, and genealogy
information. This makes the assignment of fitness through the rating
system a {\em social} activity, pursuing the objective of this work,
which is to increase engagement of the users.
\subsubsection{Graph Model} 
\subsubsection{Human Interaction Model}

\subsection{EvoDraw-Kinect}
\subsection{XYZ-Unplugged}
